---
title: "PeerAssessment1"
author: "Carlos E Rivadeneyra"
date: "January 18, 2015"
Sundar
output: html_document
---

##Reproducible research: Peer Assessment 1

###Part 1 Settings & Load and data processing 

**First default settings**

```{r}
echo = TRUE  # code visible
options(scipen = 1)  # Turn off scientific notations for numbers
```

*Load and processing data


```{r}
data <- read.csv("activity.csv")
head(data)
```

##What is mean total number of steps taken per day?

In this part we will  ignore the missing values in the dataset.

Make a histogram of the total number of steps taken each day

```{r}
library(ggplot2)
Tsteps <- tapply(data$steps, data$date, FUN = sum, na.rm = TRUE)
qplot(Tsteps, binwidth = 900, xlab = "total # steps each day")
```

```{r}
mean(Tsteps, na.rm = TRUE)
median(Tsteps, na.rm = TRUE)
```

##What is the average daily activity pattern?

```{r}
library(ggplot2)
aver <- aggregate(x = list(steps = data$steps), by = list(interval = data$interval), 
    FUN = mean, na.rm = TRUE)
ggplot(data = aver, aes(x = interval, y = steps)) + geom_line() + xlab("five minute inter") + ylab("average steps taken")
```

```{r}
aver[which.max(aver$steps), ]
```

## MISSING VALUES

1. The total number of rows with NAs:

```{r}
sum(is.na(data))
```

I will use the mean for that 5-minute interval to replace each NA value in the steps column.

```{r}
fill.value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps)) 
        filled <- c(steps) else filled <- (aver[aver$interval == interval, "steps"])
    return(filled)
}
Newdata <- data
Newdata$steps <- mapply(fill.value, Newdata$steps, Newdata$interval)
```

Histogram with the replacement of 5 minute mean interval

```{r}
TTsteps <- tapply(Newdata$steps, Newdata$date, FUN = sum)
qplot(TTsteps, binwidth = 900, xlab = "total # steps taken each day")
```


```{r}
mean(TTsteps)
```

```{r}
median(TTsteps)
```

##Are there differences in activity patterns between weekdays and weekends?

```{r}
head(Newdata)
```

```{r}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")) 
        return("weekday") else if (day %in% c("Saturday", "Sunday")) 
        return("weekend") else stop("invalid date")
}
Newdata$date <- as.Date(Newdata$date)
Newdata$day <- sapply(Newdata$date, FUN = weekday.or.weekend)
```

Lets show two graphs a time series plot of the 5 minute interval  and the average number of steps, averaged across on weekdays   weekends.

```{r}
aver1 <- aggregate(steps ~ interval + day, data = Newdata, mean)
ggplot(aver1, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) + 
    xlab("five minute interval") + ylab("# steps")
```
